---
title: "StroopProject"
author: "David Vartanian"
date: "5 September 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```
```{r dataset}
library(readr)
library(ggplot2)
library(gridExtra)
data <- read_csv("data/stroopdata.csv")
```

```{r functions}
plot_histogram <- function(x, xlabel = '', colour = I('#6688CC'), title = '', 
                           bins = 20) {
  plot <- ggplot(aes(x = x), data = data) +
    geom_histogram(aes(colour = colour), bins = bins) +
    xlab(xlabel) +
    ggtitle(title)
  return(plot)
}
```

# Stroop - Perceptual Phenomenon
### by David Vartanian

* Link to the test: https://faculty.washington.edu/chudler/java/ready.html
* Link to t-table: https://s3.amazonaws.com/udacity-hosted-downloads/t-table.jpg

### 1. What is our independent variable? What is our dependent variable?

Independent variable (predictor): Congruent and Incongruent.

Dependent variable (outcome): Time spent by people to complete each test.

### 2. What is an appropriate set of hypotheses for this task? What kind of statistical test do you expect to perform? Justify your choices.
We are trying to test whether the difference in time between taking the two 
tests is significant, using $\mu_c$ as the mean population of the *Congruent* 
test, and $\mu_i$ as the mean population of the *Incongruent* test.

$$H_0: \mu_c = \mu_i$$

The *null* hypothesis expects that the difference between $\mu_c$ and $\mu_i$
will not statistically significant.

$$H_a: \mu_c < \mu_i$$

On the other hand, the *alternative* hypothesis expects that $\mu_c$ will be 
significantly less than $\mu_i$.

Since I don't know the population standard deviation and less than 30 
observations are provided in this dataset, I will use paired T-Test this time.

### 3. Report some descriptive statistics regarding this dataset. Include at least one measure of central tendency and at least one measure of variability.

```{r echo=TRUE}
# Count:
nrow(data)
```

```{r echo=TRUE}
# Min, max, mean:
summary(data)
```

```{r echo=TRUE}
# Variance:
var(data)
# Congruent Standard Deviation
sd(data$Congruent)
# Incongruent Standard Deviation
sd(data$Incongruent)
```

### 4. Provide one or two visualizations that show the distribution of the sample data. Write one or two sentences noting what you observe about the plot or plots.

```{r}
p1 <- plot_histogram(data$Congruent, bins = 8, xlabel = 'Congruent', title = 'Congruent Distribution')
p2 <- plot_histogram(data$Incongruent, bins = 8, xlabel = 'Incongruent', title = 'Incongruent Distribution')
grid.arrange(p1, p2, ncol = 2)
```

### 5. Now, perform the statistical test and report your results. What is your confidence level and your critical statistic value? Do you reject the null hypothesis or fail to reject it? Come to a conclusion in terms of the experiment task. Did the results match up with your expectations?

```{r}
Xbar <- mean(data$Congruent)
Ybar <- mean(data$Incongruent)
N <- nrow(data)
point.estimate <- Xbar - Ybar
D <- data$Congruent - data$Incongruent
square.deviation <- (D - point.estimate)^2
variance <- sum(square.deviation) / (N - 1)
STD <- sqrt(variance)
t.statistic <- (point.estimate / (STD / sqrt(N)))
standard.error <- STD / sqrt(N)
cohens.d <- point.estimate / STD
t.critical <- 1.714  # from t-table: 23DF, alpha .95, one-tailed
ci <- t.critical * (STD / sqrt(N))
ci.upper <- point.estimate + ci
reject.null <- t.statistic > ci.upper
p.value <- 0.0001  # t=-8.020707   DF=23
```

* $\mu_c = `r Xbar`$
* $\mu_i = `r Ybar`$
* $N = `r N`$
* Point Estimate ($\mu_c - \mu_i$) = `r point.estimate`
* Variance = `r variance`
* Standard Deviation = `r STD`
* T-Statistic = `r t.statistic`
* Standard Error = `r standard.error`
* T-Critical = `r t.critical` (from t-table: 23DF, alpha .95, one-tailed)
* Confidence Interval = `r ci`
* CI - Upper = `r ci.upper`
* P-Value = `r p.value`
* Reject NULL = `r reject.null`

At the beginning it didn't match my expectations, because after doing the test
myself, I noticed a considerable difference between both congruent and 
incongruent parts. But with less than 30 observations in the dataset it's quite 
obvious that the Confidence Interval would be huge. Therefore, the T-Statistic 
would not be far enough to cross the T-Critical value in order to reject the 
NULL. In my opinion, duplicating the number of observations would be enough to 
get a more real result.

### 6. Optional: What do you think is responsible for the effects observed? Can you think of an alternative or similar task that would result in a similar effect? Some research about the problem will be helpful for thinking about these two questions!

The effect is caused by a conflict between what for the brain is easier (words)
and what we need to pay attention to (colours).

Extracted from the Stroop Effect website:

> The words themselves have a strong influence over your ability to say the 
color. The interference between the different information (what the words say 
and the color of the words) your brain receives causes a problem. There are two 
theories that may explain the Stroop effect:
>
> Speed of Processing Theory: the interference occurs because words are read 
faster than colors are named.
Selective Attention Theory: the interference occurs because naming colors 
requires more attention than reading words.
>
>More experiments to try:
>
>* Turn the words upside down or rotate them 90 degrees.
>* Turn the words "inside out."
>* Use non-color words such as "dog" or "house."
>* Use nonsense words such as "kiw" or "thoz."
>* Compare long words to short words.
>* Use emotional words such as "sad" or "happy" or "depressed" or "angry."
>* Color only half of the word or color only the first and last letter of each 
word.



