---
title: "StroopProject"
author: "David Vartanian"
date: "5 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```
```{r dataset}
library(readr)
library(ggplot2)
library(gridExtra)
data <- read_csv("data/stroopdata.csv")
```

```{r echo=FALSE functions}
plot_histogram <- function(x, xlabel = '', colour = I('#6688CC'), title = '', bins = 20) {
  plot <- ggplot(aes(x = x), data = data) +
    geom_histogram(aes(colour = colour), bins = bins) +
    xlab(xlabel) +
    ggtitle(title)
  return(plot)
}
```

# Stroop - Perceptual Phenomenon
### by David Vartanian

Link to the test: https://faculty.washington.edu/chudler/java/ready.html
Link to t-table: https://s3.amazonaws.com/udacity-hosted-downloads/t-table.jpg

### 1. What is our independent variable? What is our dependent variable?
- is this a pre-test post-test? if so, Congruent is independent, Incongruent is dependent
Pre-test Post-test Design
Independent Variable: Congruent
Dependent Variable: Incongruent

Meet specifications: Q1: Question response correctly identifies the independent and dependent variables in the experiment.

### 2. What is an appropriate set of hypotheses for this task? What kind of statistical test do you expect to perform? Justify your choices.
- null hypothesis: the time to complete the second test is not significantly different
- alternative hypothesis: the time to complete the second test is significantly higher

Meet specifications:
- Q2a: Null and alternative hypotheses are clearly stated in words and mathematically. Symbols in the mathematical statement are defined.
- Q2b: A statistical test is proposed which will distinguish the proposed hypotheses. Any assumptions made by the statistical test are addressed.

### 3. Report some descriptive statistics regarding this dataset. Include at least one measure of central tendency and at least one measure of variability.
- count, min, max on both Congruent and Incongruent variables
- mean, median times (Congruent, Incongruent)
```{r}
summary(data)
```
- variance, standard deviation
```{r}
var(data)
sd(data$Congruent)
sd(data$Incongruent)
```

Meet specifications: Q3: Descriptive statistics, including at least one measure of centrality and one measure of variability, have been computed for the dataset’s groups.

### 4. Provide one or two visualizations that show the distribution of the sample data. Write one or two sentences noting what you observe about the plot or plots.
- plot distribution

```{r}
p1 <- plot_histogram(data$Congruent, bins = 8, xlabel = 'Congruent', title = 'Congruent Distribution')
p2 <- plot_histogram(data$Incongruent, bins = 8, xlabel = 'Incongruent', title = 'Incongruent Distribution')
grid.arrange(p1, p2, ncol = 2)
```

Meet specifications: Q4: One or two visualizations have been created that show off the data, including comments on what can be observed in the plot or plots.

### 5. Now, perform the statistical test and report your results. What is your confidence level and your critical statistic value? Do you reject the null hypothesis or fail to reject it? Come to a conclusion in terms of the experiment task. Did the results match up with your expectations?
calculate:
  - sample mean difference

```{r}
Xbar <- mean(data$Congruent)
Ybar <- mean(data$Incongruent)
N <- nrow(data)
point.estimate <- Xbar - Ybar
D <- data$Congruent - data$Incongruent
square.deviation <- (D - point.estimate)^2
variance <- sum(square.deviation) / (N - 1)
STD <- sqrt(variance)
t.statistic <- (point.estimate / (STD / sqrt(N)))
standard.error <- STD / sqrt(N)
cohens.d <- point.estimate / STD
t.critical <- 1.714  # from t-table: 23DF, alpha .95, one-tailed
ci <- t.critical * (STD / sqrt(N))
ci.upper <- point.estimate + ci
reject.null <- t.statistic > ci.upper
p.value <- 0.0001  # t=-8.020707   DF=23

ggplot(aes(x = Congruent, y = Incongruent), data = data) +
  geom_point() +
  geom_smooth(method = 'lm', size = 1, se = FALSE)
```

```{r}
data2 <- rbind(data, data)
Xbar <- mean(data2$Congruent)
Ybar <- mean(data2$Incongruent)
N <- nrow(data2)
point.estimate <- Ybar - Xbar
D <- data2$Incongruent - data2$Congruent
square.deviation <- (D - point.estimate)^2
variance <- sum(square.deviation) / (N - 1)
STD <- sqrt(variance)
t.statistic <- (point.estimate / (STD / sqrt(N)))
standard.error <- STD / sqrt(N)
cohens.d <- point.estimate / STD
t.critical <- 1.676  # from t-table: 23DF, alpha .95, one-tailed
ci <- t.critical * (STD / sqrt(N))
ci.upper <- point.estimate + ci
reject.null <- t.statistic > ci.upper
p.value <- 0.0310  # t=-2.29744   DF=23

ggplot(aes(x = Congruent, y = Incongruent), data = data) +
  geom_point() +
  geom_smooth(method = 'lm', size = 1, se = FALSE)
```

  - standard deviation of the mean
  - t-critical
  - t-statistic
  - standard error
  - confidence interval
  - p-value
  - statistical significance
  - comments about statistic

> Meet specifications: Q5: A statistical test has been correctly performed and reported, including test statistic, p-value, and test result. The test results are interpreted in terms of the experimental task performed.


### 6. Optional: What do you think is responsible for the effects observed? Can you think of an alternative or similar task that would result in a similar effect? Some research about the problem will be helpful for thinking about these two questions!
make some research about the cientific reason of taking so much more time when colour words have different colours

Meet specifications: Q6: Hypotheses regarding the reasons for the effect observed are presented. An extension or related experiment to the performed Stroop task is provided, that may produce similar effects.











